Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-05-07 code/rmarkdown/article.tex
article fixes applied
Former-commit-id: 3a40abac5ca3a8656f5c5f1ed1fde5f7d46f5b73

We successfully mapped the main spatial patterns of adaptive (if
present) and neutral variation for each of the 26 species. The outlier
detection methods detected adaptive genetic variation in 4 species.
Generally, two outlier detection methods gave fairly consistent results
(mean 87.26 \% loci per species 4.97 SD). Of the species that were
associated with adaptive genetic variation, only a few loci were
detected as under selection (mean 1.77 \% loci per species 1.77 SD).
After identifying the loci under selection, we used NMDS analyses to
identify the main gradients of adaptive (if detected) and neutral
variation for each species. Generally, only a couple of continuous
dimensions were needed to sufficiently describe the adaptive genetic
variation for each species (minimum \(K\) NA; mean \(K\) 2; max \(K\) 2;
Table S2). On the other hand, a few more continuous dimensions were
often needed for neutral genetic variation for each species (minimum
\(K\) 2; mean \(K\) 2.4230769; max \(K\) 4; Table S2). The ordinations
resulting from the NMDS analyses were used to construct an adaptive (if
present) and neutral genetic space for each species. The spatial
distribution of these genetic spaces (Figures S29--S54) generally show
patterns of spatial autocorrelation--especially the adaptive genetic
spaces (eg. \emph{Phyteuma hemisphaericum}; S49). Additionally, most
species show east-west patterns in their intra-specific genetic
variation (eg. \emph{Dryas octopetala}; S36).

The dissimilarity matrices generated using the surrogate variables
correlated with the dissimilarity matrices based on the genetic
characteristics of individuals for most species (Table 1). Specifically,
environmental variation was found to significantly correlate with
adaptive genetic variation for 1 species (NA \% associated with loci
under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
\(\pm\) NA SD for significant models). Additionally, geographic
distances between planning units were found to significantly correlate
with neutral genetic variation among planning units for 25 species
(96.15 \% of all the species investigated;\emph{P} \textless{} 0.05;
mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
These results suggest that prioritizations with different environmental
characteristics tend to have individuals with different adaptive genetic
characteristics. Similarly, planning units that are further apart tend
to have individuals with different neutral genetic characteristics.
After verifying that the geographic and environmental variables
correlate with genetic variation, the next step was to determine how
effective these variables were as surrogates in the context of
conservation planning. The dissimilarity matrices generated using the
surrogate variables correlated with the dissimilarity matrices based on
the genetic characteristics of individuals for most species (Table 1).
Specifically, environmental variation was found to significantly
correlate with adaptive genetic variation for 1 species (NA \%
associated with loci under selection; \emph{P} \textless{} 0.05; mean
marginal 0.02 \(R^2\) \(\pm\) NA SD for significant models).
Additionally, geographic distances between planning units were found to
significantly correlate with neutral genetic variation among planning
units for 25 species (96.15 \% of all the species investigated;\emph{P}
\textless{} 0.05; mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for
significant models). These results suggest that prioritizations with
different environmental characteristics tend to have individuals with
different adaptive genetic characteristics. Similarly, planning units
that are further apart tend to have individuals with different neutral
genetic characteristics. After verifying that the geographic and
environmental variables correlate with genetic variation, the next step
was to determine how effective these variables were as surrogates in the
context of conservation planning.

The environmental and geographic variables were moderately effective
surrogates for genetic variation in most species (Figure 2). The
relationship between the proportion of adaptive genetic variation
secured in a prioritization and the proportion environment variation it
also secured differed depending on the species for which the
prioritization was generated
(\(\chi^2_{4933.615872} P = 2.7275469\times 10^{-11}\)). The slope for
each species were all significantly different to zero (minimum
\(z value = 31.6562515\); maximum
\(P = r `max(env.surrogate.spp.DF\)p.value)`\$; Table S3). The
explanatory power of the environmental surrogates were also generally
quite high for most species (mean 0.5746939 \(R^2\) 0.0365087 SD; Figure
S56), with exception to \emph{Carex sempervirens}. The nature of these
relationships suggest that most of the adaptive variation for a given
species can be secured in a prioritization using moderately high
environmental-based targets. Similarly, the proportion of neutral
genetic variation in a prioritization was affected by the interaction
between species and the proportion of neutral variation secured in the
prioritization (\(\chi^2_{4.462023\times 10^{4}} P = 0\)). The slope for
each species were also significantly different to zero (minimum
\(z value = 23.3822268\); maximum \$P = 0; Table S3). Generally, the
explanatory power of geographic variables were higher than that of
environmental variables (mean 0.5495176 \(R^2\) 0.1272203 SD; Figure
S57). Compared to the environmental variables, the slopes of the
relationships suggest that very high geographic-based targets are
required to generate prioritizations that secure a sufficient proportion
of neutral genetic variation for most species. Additionally, all species
were associated with prioritizations that secured a large proportion of
environmental or geographic variation (\textgreater{} 70 \%), yet
preserved only a very small proportion of the genetic variation
(\textless{} 10 \%). These results suggest while prioritizations
generated using surrogate-based targets tend to secure more genetic
variation, prioritizations generated using just these targets can result
in poor solutions.

To determine if the addition of environmental and geographic targets
could improve the effectiveness of prioritizations in a more realistic
context, we generated prioritizations according to three different
scenarios (Figure 3). The prioritizations generated under the
single-species scenario showed relatively homogeneous selection
frequencies across the study area--regardless of the combination of
targets used to generate them (Figures 3a, 3b, 3c). On the other hand,
the prioritizations generated under the multi-species scenario showed
different selection frequencies depending on the targets used to
generate them. The prioritization generated using only amount-based
targets contained fewer planning units (Figure 3d; \(n = 14\)) than the
prioritizations generated using additional of surrogate-based targets
(Figure 3e; \(n = 33\)) or genetic-based targets (Figure 3f;
\(n = 27\)). The prioritizations generated under the multi-species with
cost scenario had different spatial distributions to the previous two
scenarios. These prioritizations tended to concentrate conservation
effort towards areas with higher elevation. Similar to the second
scenario, the amount-based prioritization contained less panning units
(Figure 3g; \(n = 14\)) than either the surrogate-based (Figure 3h;
\(n = 48\)) or the genetic-based prioritizations ((Figure 3i;
\(n = 34\))).

@@

The prioritizations generated using additional surrogate-based targets
met the neutral genetic variation targets for more species than the
prioritizations generated using just amount-based targets (Figure 3).
The biggest difference between the prioritizations generated using
additional surrogate-based targets and just the amount-based targets was
in the single-species scenarios (\texttt{r} \% and \texttt{r} \% of
species meeting genetic variation targets respectively). In the other
two scenarios, the amount-based prioritizations tended to perform
slightly better (\texttt{r} \% of species with targets met), but were
outperformed by prioritizations with additional surrogate-based targets.
Although the surrogate-based targets tended to fulfill the neutral
genetic variation targets for most species, they captured only a very
small proportion of neutral genetic variation for two species in
particular (\emph{Gypsophila\_repens} and \emph{Luzula\_alpinopilosa}).
These results suggest that surrogate-based targets can substantially
improve the proportion of neutral genetic variation secured in a
prioritization.

>>>>

We successfully mapped the main spatial patterns of adaptive (if
present) and neutral variation for each of the 26 species. The outlier
detection methods detected adaptive genetic variation in four species.
Generally, two outlier detection methods gave fairly consistent results
(mean 87.26 \% loci per species \(\pm\) 4.97 SD). Of the species that
were associated with adaptive genetic variation, only a few loci were
detected as under selection (mean 1.77 \% loci per species \(\pm\) 1.77
SD). After identifying the loci under selection, we used NMDS analyses
to identify the main gradients of adaptive (if detected) and neutral
variation for each species. Generally, only a couple of continuous
dimensions were needed to sufficiently describe the adaptive genetic
variation for each species (all species \(K = 2\); Table S2). On the
other hand, a few more continuous dimensions were often needed for
neutral genetic variation for each species (minimum \(K = 2\); mean
\(K = 2.4230769\); max \(K = 4\); Table S2). The ordinations resulting
from the NMDS analyses were used to construct an adaptive (if present)
and neutral genetic space for each species. The spatial distribution of
these genetic spaces (Figures S29--S54) generally show patterns of
spatial autocorrelation--especially the adaptive genetic spaces (eg.
\emph{Phyteuma hemisphaericum}; S49). Additionally, most species show
east-west patterns in their intra-specific genetic variation (eg.
\emph{Dryas octopetala}; S36).

The dissimilarity matrices generated using the surrogate variables
correlated with the dissimilarity matrices based on the genetic
characteristics of individuals for most species (Table 1). Specifically,
environmental variation was found to significantly correlate with
adaptive genetic variation for one species (25 \% associated with loci
under selection; \emph{P} \textless{} 0.05; mean marginal 0.02 \(R^2\)
\(\pm\) NA SD for significant models). Additionally, geographic
distances between planning units were found to significantly correlate
with neutral genetic variation among planning units for 25 species
(96.15 \% of all the species investigated; \emph{P} \textless{} 0.05;
mean marginal 0.21 \(R^2\) \(\pm\) 0.17 SD for significant models).
These results suggest that prioritizations with different environmental
characteristics tend to have individuals with different adaptive genetic
characteristics. Similarly, planning units that are further apart tend
to have individuals with different neutral genetic characteristics.
After verifying that the geographic and environmental variables
correlate with genetic variation, the next step was to determine how
effective these variables were as surrogates in the context of
conservation planning.

The environmental and geographic variables were moderately effective
surrogates for genetic variation in most species (Figure 2). The
relationship between the proportion of adaptive genetic variation
secured in a prioritization and the proportion environment variation it
also secured differed depending on the species for which the
prioritization was generated
(\(\chi^2_{3} = {4933.62} P = pvalString(2.7275469\times 10^{-11})\)).
The slope for each species were all significantly different to zero
(minimum \textit{z} = 31.66; maximum \(P = < 0.001\); Table S3). The
explanatory power of the environmental surrogates were also generally
quite high for most species (mean 0.57 \(R^2 \pm\) 0.04 SD; Figure S56),
with exception to \emph{Carex sempervirens}. The nature of these
relationships suggest that most of the adaptive variation for a given
species can be secured in a prioritization using moderately high
environmental-based targets. Similarly, the proportion of neutral
genetic variation in a prioritization was affected by the interaction
between species and the proportion of neutral variation secured in the
prioritization (\(\chi^2_{25} = 4.462023\times 10^{4}\); \(P\) =
\textless{} 0.001). The slope for each species were also significantly
different to zero (minimum \(z value = 23.3822268\); maximum \$P = 0;
Table S3). Generally, the explanatory power of geographic variables were
higher than that of environmental variables (mean 0.55 \(R^2 \pm\) 0.13
SD; Figure S57). Compared to the environmental variables, the slopes of
the relationships suggest that very high geographic-based targets are
required to generate prioritizations that secure a sufficient proportion
of neutral genetic variation for most species. Additionally, all species
were associated with prioritizations that secured a large proportion of
environmental or geographic variation (\textgreater{} 70 \%), yet
preserved only a very small proportion of the genetic variation
(\textless{} 10 \%). These results suggest while prioritizations
generated using surrogate-based targets tend to secure more genetic
variation, prioritizations generated using just these targets can result
in poor solutions.

To determine if the addition of environmental and geographic targets
could improve the effectiveness of prioritizations in a more realistic
context, we generated prioritizations according to three different
scenarios (Figure 3). The prioritizations generated under the
single-species scenario showed relatively homogeneous selection
frequencies across the study area--regardless of the combination of
targets used to generate them (Figures 3a, 3b, 3c). On the other hand,
the prioritizations generated under the multi-species scenario showed
different selection frequencies depending on the targets used to
generate them. The prioritization generated using only amount-based
targets contained fewer planning units (Figure 3d; \(n = 14\)) than the
prioritizations generated using additional of surrogate-based targets
(Figure 3e; \(n = 33\)) or genetic-based targets (Figure 3f;
\(n = 27\)). The prioritizations generated under the multi-species with
cost scenario had different spatial distributions to the previous two
scenarios. These prioritizations tended to concentrate conservation
effort towards areas with higher elevation. Similar to the second
scenario, the amount-based prioritization contained less panning units
(Figure 3g; \(n = 14\)) than either the surrogate-based (Figure 3h;
\(n = 48\)) or the genetic-based prioritizations (Figure 3i;
\(n = 34\)).

@@

The prioritizations generated using additional surrogate-based targets
met the neutral genetic variation targets for more species than the
prioritizations generated using just amount-based targets (Figure 3).
The biggest difference between the prioritizations generated using
additional surrogate-based targets and just the amount-based targets was
in the single-species scenarios (\texttt{r} \% and \texttt{r} \% of
species meeting genetic variation targets respectively). In the other
two scenarios, the amount-based prioritizations tended to perform
slightly better (\texttt{r} \% of species with targets met), but were
outperformed by prioritizations with additional surrogate-based targets.
Although the surrogate-based targets tended to fulfill the neutral
genetic variation targets for most species, they captured only a very
small proportion of neutral genetic variation for two species in
particular (\emph{Gypsophila repens} and \emph{Luzula alpinopilosa}).
These results suggest that surrogate-based targets can substantially
improve the proportion of neutral genetic variation secured in a
prioritization.
